---
title: "post02-vibha-seshadri"
author: "Vibha Seshadri"
date: "November 20, 2017"
output: html_document
---
## Stringr and Regex: A Force to Reckon With

#### Introduction
I'm just going to start off by saying that I am a **BIG** *Harry Potter* fan. My favorite book in the series was  *Harry Potter and the Deathly Hallows*! I loved how well J.K. Rowling tied everything in the series together with so much attention to detail and depth! That's why in this blog post, I am going to explain how we can use the methods in the `stringr` package and regex to analyze the master piece that is the text of the seventh Harry Potter book! Let's begin! 

#### What exactly is regex?
[According to the regex rdocumentation](https://www.rdocumentation.org/packages/base/versions/3.4.1/topics/regex), regex stands for "Regular Expression" and "speficies a pattern which describes a set of strings". Regex has specific syntax that is used in creating patterns for the computer to interpret and match. Take a look at [this cheatsheet](http://regexlib.com/CheatSheet.aspx?AspxAutoDetectCookieSupport=1) to learn more about regex syntax! There are many uses of Regex! [According to Rex Egg](http://www.rexegg.com/regex-uses.html), a website that contains many regex tutorials, Some examples include: getting text from a file, validating input from users into forms, searching for and replacing pages of code, using the search features of creativity software, renaming many files, and much more. 

#### What exactly is `stringr`?
`stringr` is a package in `R` that allows for easy string manipulation and pattern matching in character vectors.
[According to the cran.r webpage on stringr](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html), `stringr` supports four types of operations: character manipulations, whitespace manipulations, locale operations (ie changing case or sorting a vector), and pattern matching functions. [According to the stringr rdocumentation](https://www.rdocumentation.org/packages/stringr/versions/1.1.0), `stringr` was created by Hadley Wickham and provides a consistent manner in working with strings because "all functions and argument names are consistent, and all functions deal with missing values in the same way." This is what allows `stringr` to suppor the piping operator (`%>%`) and use the output of one function as the input to another function.

`stringr` functions are useful because they are very versatile yet conistent. They can be used to clean and manipulate text as well as find patterns in text when combined with regex. We will see all of this in the example with the text of the seventh Harry Potter book.

#### Motivation
So exactly why am I explaining how to use `stringr` to analyze this text? Well let's just say that I want to save you the trouble of physically analyzing words and patterns within a text file the next time you analyze text (because, that's what we do for fun, right?). Okay jokes aside, it's no secret that anyone who loves reading pays attention to the author's word choice, sentence structure, other semantic and grammar choices, and analyzes what messages those choices convey to the readers. Well the beauty of `stringr` and regex is that is allows us to pay attention to these same things, but in more detail since the combination of `stringr` and regex can go through an entire text file and find what we are looking for in a fraction of the time it would take a human to do the same thing! Using the string manipulation methods in `stringr` in combination with regex can be used in many situations, such as replacing many instances of the same word or phrase in one text file or finding a repeating pattern in a text file, without looping through every line in the text file. In R, you may be able to get away with using some vectorized operations, but in other programming languages you won't be able to do this! Luckily, regex is universal across programming languages and is a handy skill to learn to avoid looping through long pieces of text. At this point some of you may be wondering what's so bad about using loop operations or parse an entire file to do string manipulation and pattern matching? Well, [according to a post by Oana Platon on the Microsoft Developer Blog](https://blogs.msdn.microsoft.com/oanapl/2009/04/04/performance-comparison-regex-versus-string-operations/) that would work, but it's a lot more work and there's a lot more room for error. For instance, if you were trying to find every word that started with the letters "app" and then return the maximum length of a word that started with "app" you could use a loop to find these words and check if the substring from the first to third index of the words matched "app", and then you use the max function on a list of the length of such words, and return the maximum value. But what if you could skipping looking at each word individually all together? And what if you could avoid errors that arise when comparing strings that are of different lengths? This is where regex makes things easy. Consider the following example.

```{r}
# create example vector
ex <- c("application", "broke", "cake", "", "appelate", "eat,", "apple", "apothecary", "table", "append", "thanksgiving", "apostle", "ape", "appease")

# find all the words in ex that start with app using regex
# pattern ^app and stringr function str_detect
app <- ex[str_detect(ex, "^app")]

# get their length using stringr function str_length
len <- str_length(app)

# return the max length
max(len)
```

We didn't have to loop or look at any of the elements in the character vector `ex` individually! Imagine if `ex` was as long as the text of the seventh Harry Potter book and had `r length(hp_seven)` words or `r sum(nchar(hp_seven))` characters! We would not want to look at any word individually ourselves, but rather have words that match the description we are looking for returned to us easily. That is what regex and `stringr` functions allow us to do.

#### Examples of stringr in action!
```{r}
# load stringr package
library(stringr)

# concatenate the url to get the text file
beg <- "https://ia801309.us.archive.org/15/items/Book5TheOrderOfThePhoenix/"
end <- "Book%207%20-%20The%20Deathly%20Hallows_djvu.txt"
url <- paste0(beg, end)

# read in the text file from the url
hp_seven <- readLines(url)

# get rid of lines that indicate the page number and title, as they are not 
# directly from the novel
hp_seven <- hp_seven[!str_detect(hp_seven, "Page \\| [1-9]*[0-9]")]
hp_seven <- hp_seven[!str_detect(hp_seven, 
              "Harry Potter and the Deathly Hallows - J.K. Rowling ")]

# split up the lines into words
hp_seven <- str_split(hp_seven, " ")
hp_seven <- unlist(hp_seven)

# get rid of entries in hp_seven that are empty strings
hp_seven <- hp_seven[str_detect(hp_seven, "")]
```

#### Conclusion
#### References
- [Where I got the Harry Potter book 7 text file from](https://archive.org/stream/Book5TheOrderOfThePhoenix/Book%207%20-%20The%20Deathly%20Hallows#page/n15/mode/2up)
- [Regex v. string operations](https://blogs.msdn.microsoft.com/oanapl/2009/04/04/performance-comparison-regex-versus-string-operations/)
- [Regex rdocumentation](https://www.rdocumentation.org/packages/base/versions/3.4.1/topics/regex)
- [Regex cheatsheet](http://regexlib.com/CheatSheet.aspx?AspxAutoDetectCookieSupport=1)
- [Uses of Regex](http://www.rexegg.com/regex-uses.html)
- [stringr rdocumentation](https://www.rdocumentation.org/packages/stringr/versions/1.1.0),
- [stringr documentation](https://cran.r-project.org/web/packages/stringr/stringr.pdf)

- [Regex and stringr basics](http://stringr.tidyverse.org/articles/regular-expressions.html)
