---
title: 'Homework 1: Summarizing Variables'
author: "Vibha Seshadri"
date: "September 18, 2017"
output: github_document
---

```{r loadVars}
# loading data
load("../data/nba2017-salary-points.RData")

# listing available objects
ls()
```

```{r simplifyMillions}
# simplifying representation of dollar quantities stored in salary
milSalary <- salary/1000000
milSalary <- round(milSalary, digits = 2)
```

```{r expAsInt}
# creating a new vector that holds numeric representation of experience, previously a vector of type char
experience <- replace(experience, experience == "R", "0")
experience <- as.integer(experience)
```

```{r posFactor}
# changing variable position from character vector to factor with more descriptive labels
position <- factor(position, labels = c("center", "power_fwd", "point_guard", "small_fwd", "shoot_guard"))

# calculate frequency of positions in factor, position
table(position)
```

```{r scatterPlot}
# creating a scatterplot to analyze the relationship between points and salary
plot(points, milSalary, xlab = "Points", ylab = "Salary (millions)", 
     main = "Relationship Between Points and Salary", col = c("light blue"),
     cex.main = "1", cex.lab = "1", pch = 16)
```

The scatterplot above illustrates a positive relationship between the number of `points` a player makes and his resulting `salary`. The majority of the data is clustered in the lower left corner of the scatterplot, and begins to uncluster and grow with a positive slope as players gain more than `r median(points)` total points. There are, however, many points on the graph which do not follow the positive trend of the majority of points. The most total points a player has made is `r max(points)` points, but this player does not have the highest salary. This player with `r max(points)` has a salary of $`r milSalary[which.max(points)]`. The highest salary appears to be around $`r max(milSalary)` dollars and is the salary of a player who has a total of `r points[which.max(salary)]` points.