---
title: 'Homework 1: Summarizing Variables'
author: "Vibha Seshadri"
date: "September 18, 2017"
output: github_document
---

```{r loadVars}
# loading data
load("../data/nba2017-salary-points.RData")

# listing available objects
ls()
```

####A bit of data preprocessing
```{r simplifyMillions}
# simplifying representation of dollar quantities stored in salary
milSalary <- salary/1000000
milSalary <- round(milSalary, digits = 2)
```

```{r expAsInt}
# creating a new vector that holds numeric representation of experience, previously a vector of type char
experience <- replace(experience, experience == "R", "0")
experience <- as.integer(experience)
```

```{r posFactor}
# changing variable position from character vector to factor with more descriptive labels
position <- factor(position, labels = c("center", "power_fwd", "point_guard", "small_fwd", "shoot_guard"))

# calculate frequency of positions in factor, position
table(position)
```

####Scatterplot of `points` and `salary`
```{r scatterPlot}
# creating a scatterplot to analyze the relationship between points and salary
plot(points, milSalary, xlab = "Points", ylab = "Salary (millions)", 
     main = "Relationship Between Points and Salary", col = c("light blue"),
     cex.main = "1", cex.lab = "1", pch = 16)
```

The scatterplot above illustrates a positive relationship between the number of `points` a player makes and his resulting `salary`. The majority of the data is clustered in the lower left corner of the scatterplot, and begins to uncluster and grow with a positive slope as players gain more than `r median(points)` total points. There are, however, many points on the graph which do not follow the positive trend of the majority of points. The most total points a player has made is `r max(points)` points, but this player does not have the highest salary. This player with `r max(points)` has a salary of $`r milSalary[which.max(points)]`. The highest salary appears to be around $`r max(milSalary)` dollars and is the salary of a player who has a total of `r points[which.max(salary)]` points.

####Correlation between `points` and `salary`
```{r manualCalc}
# Calculating number of individuals
n <- length(player)
n

# Calculating the mean of variable X (points)
meanX <- (1/n) * sum(points)
meanX

# Calculating the mean of Y (salary)
meanY <- (1/n) * sum(salary)
meanY

# Calculating variance of X (points)
varX <- (1/(n-1)) * sum((points - meanX) ^ 2)
varX

# Calculating variance of Y (salary)
varY <- (1/(n-1)) * sum((salary - meanY) ^ 2)
varY

# Standard deviation of X (points)
sdX <- (varX) ^ (1/2)
sdX

# Standard deviation of Y (salary)
sdY <- (varY) ^ (1/2)
sdY

# Covariance between X and Y
covXY <- (1/(n-1)) * sum((points - meanX) * (salary-meanY))
covXY

# Correlation between X and Y
corXY <- covXY/(sdY * sdX)
corXY
```

####Simple Linear Regression
```{r linRegress}
# Calculating b1, slope
b1 <- corXY * (sdY/sdX)

# Calculating b0, intercept
b0 <- meanY - b1 * meanX

# Calculating y-hat, predicted salary
yHat <- b0 + b1 * points

# Summary statistics of vector yHat
summary(yHat)
```

The above code chunk contains the regression equation `ŷ = b0 + b1X`. You can interpret the slope, `b1`, as the amount, in dollars, a player's salary will increase for each point they gain in their career in the NBA. You can interpret the intercept, `b0`, as the amount a player makes in dollars when they begin their career, or when they have earned 0 points total in the NBA.

```{r predictSalary}
# Predicted salary for player with 0 points
b0 + b1 * 0

# Predicted salary for player with 100 points
b0 + b1 * 100

# Predicted salary for player with 500 points
b0 + b1 * 500

# Predicted salary for player with 1000 points
b0 + b1 * 1000

# Predicted salary for player with 2000 points
b0 + b1 * 2000
```

####Plotting the regression line
```{r plotRegress}
pReg <- plot(points, milSalary, xlab = "Points", ylab = "Salary (millions)", 
     main = "Relationship Between Points and Salary", col = c("light blue"),
     cex.main = "1", cex.lab = "1", pch = 16)
# reg <- lm(points ~ salary, data = yHat)
# coeff = coefficients(reg)
# abline(reg, a = b0, b = b1, col = c("black"), lwd = 2)
abline(a = b0, b = b1, col = c("black"), lwd = 2)
```

